---
title: "MSDS 6372 Project 2"
output: 
    html_document:
      keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readxl)
library(dplyr)
library(plyr)
library(glmnet)
library(ISLR)
library(leaps)
library(ggplot2)
library(ROCR)
library(MASS)
library(pheatmap)
library(randomForest)
library(gmodels)
library(vcd)
library(Amelia)
library(e1071)
library(corrplot)
library(rpart)
library(ROSE)
library(caret)

usePackage <- function(p) {

  if (!is.element(p, installed.packages()[,1]))

    install.packages(p, dep = TRUE)

  require(p, character.only = TRUE)

}
```
##Data Set 1: Osteoporosis in Women

From Hosmer, Lemeshow, and Sturdivant (2013), Applied Logistic Regression, 3rd Edition.
The Global Longitudinal Study of Osteoporosis in Women (GLOW) is an international study of osteoporosis in women aged 55 years and over. The major goals of the study are to examine prevention and treatment of fractures and distribution of risk factors among older women. Complete details on the study as well as a list of GLOW publications may be found at the Center for Outcomes Research web site, <http://www.outcomes-umassmed.org/glow>. 
There are over 60K observations in the original data set. This data set contains a sample of 500 of them. 
The link below is to a website with the data set and description of the variables. The data set in question is called "glow500".

<https://www.umass.edu/statdata/statdata/data/glow/index.html>
Note: If you choose this data set, you MAY NOT use the Hosmer, Lemeshow, and Sturdivant text to help you in your analysis.
You may only use Chapter 1 in order to obtain a description of the data.

Of course if you dont have the book

<https://www.umass.edu/statdata/statdata/data/glow/glow.pdf>
provides definitions to the variables.


The Global Longitudinal Study of Osteoporosis in Women (GLOW) (2005-2014) was a prospective cohort study of physician practices in the provision of prophylaxis and treatment against osteoporotic fractures. The goal of this research was to improve understanding of the risk and prevention of osteoporosis-related fractures among female residents of 10 countries who were 55 years of age and older. GLOW enrolled over 60,000 women through over 700 physicians in 10 countries, and conducted annual follow-up for up to 5 years through annual patient questionnaires.

#Setup: 
## Data Import and Cleaning
<p>Missing values were not detected in dataset. Special characters were removed from column headings.
<br>What we know/don't know about the sample (500): 
1. We do not know if the subjects are distributed equally around the world. We will assume that the same percentage from each region was selected for the sample in this dataset.
2. Based on the Sub_ID(Subject ID), we can assume that the datat is independent sample of participants.
3. 
</p>
```{r}

dataset <- read.csv("C:/Users/carol/OneDrive/Documents/MSDS6372/Proj2/glow500.csv", sep=",", stringsAsFactors = TRUE, header=TRUE,na.strings=c(""))

# List rows of data that have missing values 
Missing_values <- dataset[!complete.cases(dataset),]

# Create new dataset without missing data 
dataset <- na.omit(dataset)

#remove FRACSCORE feature per professor Turner
drops <- c("FRACSCORE")
dataset <- dataset[ , !(names(dataset) %in% drops)]

#Cleanup column names
colnames(dataset)[colnames(dataset)=="ï..SUB_ID"] <- "SUB_ID"



str(dataset)
```
#  Exploratory Data Analysis
## Checking the Balance of the Full dataset
<p>The current sample dataset containes a larger propotion of subjects that did not develop fracture. Building a model against this dataset could produce bias towards the majority class. 
Below you will see how many subjects with(1)/without(0) Fractures as well as the proportion percentage for each. After splitting the dataset into training and validation(test) sets, we noticed the proportion of the training and test was not any better.
<p>We fit a logistic model on the unbalanced training dataset with a threshold of .05. It shows a Precision of 1 which says there are no false positives. Recall equals 0.20 is low and indicates that we have higher number of false negatives. The F equals 0.20 is also low and suggests weak accuracy of this model.
<p>We also plotted a ROC curve to visualize the model. The AUC equals 0.764 which is low and shows the data is not balanced.
<p>We will attempt to balance the dataset in order to create a more balanced distribution of and  a better prediction.
```{r}

table(dataset$FRACTURE) 
prop.table(table(dataset$FRACTURE))

# split the data into training and validation sets
set.seed(84)
validation_index = createDataPartition(dataset$FRACTURE, p=0.75, list=FALSE)
validationData = dataset[-validation_index,c(4:14)]
trainingData = dataset[validation_index,c(4:14)]
prop.table(table(validationData$FRACTURE))
prop.table(table(trainingData$FRACTURE))

#fit a logistic regressio to unblanced training set
fit.dataset <- glm(formula=FRACTURE~ ., data = trainingData, family="binomial")
pred.fit.dataset <- predict(fit.dataset, newdata = validationData, type="response")
#Check Accuracy of fitted model.
accuracy.meas(validationData$FRACTURE,pred.fit.dataset, threshold=.05)

#Check Accuracy of Test dataset using ROC curve
roc.curve(validationData$FRACTURE, pred.fit.dataset, plotit = TRUE)
```
## Grouping Variables as Continuous, Categorical, and ID
```{r}
numericVar <- dataset[,5:8]
ID_var <- dataset[,c(1:3)]
set_noID <- dataset[4:14]
categoricalVar <- set_noID[,-c(2:5)]
```

##Create a vector of all categorical variables and run frequency 2X2s with Mosaic plots.
<p>Chi-Square Test
<br>For the 2-way tables the chisq test independence will show if 2 categorical variables are related in some population. 
<br>Null Hypothesis: The two categorical variables are independent.
<br>Alternative Hypothesis: The two categorical variables are dependent

<br>Variable: PRIORFRAC
41% of subjects with Prior Franctures also had current Fractures but only make up 25% of the overall subjects in the sample that had prior fractures. The Chi-squared p-value favors overwhemingly the alternative hypothesis that the PRIORFRAC variable is dependent on Fracture variable.

<br>Variable: PREMENO
80% of the sample subjects are not in Pre-Menopausehad of which 24% had fractures. The same frequency of 25% Premenopausal women had fractures. The Chi-squared p-value favors the null hypothesis that the PREMENO variable is independent on Fracture variable.

<br>Variable: MOMFRAC
13% of subjects have Mothers with a history of fractures. Out of those 13%, 36% of subjects also had fractures. The Chi-squared p-value favors the alternative hypothesis that the MOMFRAC variable is probably dependent on Fracture variable.

<br>Variable: ARMASSIST
62% (312/500) subjects do not have Armassist of which 20% had fractures. Of those with Armassist, 33% had fractures. The Chi-squared p-value favors the alternative hypothesis that the ARMASSIST variable is most likely dependent on Fracture variable.

<br>Variable: SMOKE
In the dataset, 93% of subjects are non-smokers of which 26% had fractures. 7% of the subjects who were smokers of which 26% had no fractures. Although the subjects are not balance in smoker vs non-smoker category, the p-value for Chi-squared test shows .47 we favor the alternative hypothesis that the Smoke variable is dependent on the Fracture. 

<br>Variable: RATERISK 
Raterisk shows the frequency of subjects in each Raterisk level is between 29%-33%. This is pretty even in terms of how many subjects are within each Raterisk. For those that did have Fractures, their probability of a fracture increased with the level of Raterisk. This makes sense. </p>
```{r}
categoricalVarVec  <- c("PRIORFRAC","PREMENO","MOMFRAC","ARMASSIST","SMOKE","RATERISK")
for(categoricalVar in categoricalVarVec){ 
  CrossTable(dataset[,categoricalVar], dataset$FRACTURE, chisq = TRUE , expected = TRUE, dnn=c(categoricalVar,"FRACTURE")) 
  mosaicplot(CrossTable(dataset[ ,categoricalVar], dataset$FRACTURE)$t, main=paste("FRACTURE vs",categoricalVar, sep=" "), xlab=categoricalVar, ylab= "FRACTURE", color=T)
}
```
## Summary Tables
<p>Assumptions
This is a prospective study which means its a study over time of a group of similar individuals who differ with respect to certain factors under a study and how these factors affect rates of a certain outcome (Fracture vs No-Fracture)
Linearity - 
Independence of errors - Based on SUB_ID(Subject ID) we confirm each record is an independent sample.
Multicollinearity - Weight and BMI are highly correlated but we will remove one from the model.
```{r fig.width=10}
# display the first 20 rows
print(head(dataset, n=20))

# display the dimensions of the dataset
print(dim(dataset))

# list types for each attribute
print(sapply(dataset,class))

# summarize the dataset
print(summary(set_noID))

# Standard Deviations for the non-categorical columns
std=sapply(set_noID,sd) 
print('The standard deviations are:')
print(std)

# Skewness
#The further the distribution of the skew value from zero, 
# the larger the skew to the left (negative skew value) or right (positive skew value).
#library(e1071) # the library for skewness
library(e1071)

skew=apply(numericVar, 2, skewness) 
print(skew)

# Correlations
library(corrplot)

#Full dataset without ID columns
corrplot(cor(set_noID), method = "number", type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45,main="Correlation - Full Dataset")

#Training data set without ID columns
#split the data into training and validation sets
library(caret)

set.seed(84)
validation_index = createDataPartition(dataset$FRACTURE, p=0.75, list=FALSE)
validationData = set_noID[-validation_index,]
trainingData = set_noID[validation_index,]

corrplot(cor(trainingData), method = "number", type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45, main="Correlation - Training Dataset")


# Data visualizations
dataset_numeric = numericVar

#Histograms
par(mfrow=c(3,4)) # put four figures in a row (2*4)
for (i in 1:4) {
  hist(dataset_numeric[,i],main=names(dataset_numeric)[i])
}

#Density Plots
par(mfrow=c(3,4))
for(i in 1:4) {
  plot(density(dataset_numeric[,i]), main=names(dataset_numeric)[i])
}

#Box And Whisker Plots
par(mfrow=c(3,4))
for(i in 1:4) {
  boxplot(dataset_numeric[,i], main=names(dataset_numeric)[i])
}

#Barplots, which is used to count the accurances for categorical attributes 
dataset_categorical = set_noID[,-c(2:5)]
par(mfrow=c(1,3))
for(i in 1:6) {
  counts <- table(dataset_categorical[,i]) # get the count for each categorical value
  name <- names(dataset_categorical)[i]
  barplot(counts, main=name,col=c("lightblue","darkred"))
}

#Multivariate Visualization 
library(corrplot) # for function corrplot()
correlations1=cor(dataset_numeric)
print(correlations1)
par(mfrow=c(1,1))
corrplot(correlations1, methods="circle")

# pair-wise scatterplots of the numeric attributes
par(mfrow=c(1,1))
pairs(dataset_numeric)

#Scatterplot Matrix By Class (use different color to distinguish different class)
par(mfrow=c(1,1))
pairs(dataset_numeric, col=dataset[,5])

#Density By Class
library(caret)

# load the data
data(iris)

# density plots for each attribute by class value
x <- dataset_numeric
y <- dataset[,5]
scales <- list(x=list(relation="free"), y=list(relation="free"))
par(mfrow=c(1,1))
featurePlot(x=dataset_numeric, y=dataset[,5], plot="density", scales=scales)
#Box And Whisker Plots By Class
featurePlot(x=dataset_numeric, y=dataset[,5], plot="box")
```
#Logistic Regression
## Train / Test
<p>Training set will be 70% of dataset and Test set will be remaining 30%</p>
```{r}
#smp_size <- floor(0.70 * nrow(dataset))
#set.seed(1234)
#train_ind <-sample(seq_len(nrow(dataset)), size=smp_size)
#test <-dataset[-train_ind,]
#train <-dataset[train_ind,]



# split the data into training and validation sets
library(caret)
set.seed(84)
validation_index = createDataPartition(dataset$FRACTURE, p=0.75, list=FALSE)
validationData = set_noID[-validation_index,]
trainingData = set_noID[validation_index,]

#check for Missing Data
sapply(trainingData,function(x) sum(is.na(x)))
sapply(trainingData, function(x) length(unique(x)))
missmap(trainingData, main = "Missing values vs observed") #library(Amelia)

```
##Build Model
<p>Question of Interest?
What are the odds of getting a fracture, given certain conditions?</p> 
```{r}
set.seed(84)
model <- glm(FRACTURE~.,family = "binomial" (link='logit'), data=set_noID)
model
summary(model)
```
<p>Interpretation of logistic regression model:
Weight, height, BMI, Premeno, Armassist, and Smoke are not statistically significant variables. Priorfrac and Age are statistically significant variables and have the lowest p-value indicating a strong association with having a Fracture. 
```{r}
##glmnet
dat <- categoricalVar

#Get Training Set
dat.train <- set_noID

dat.train.x <- dat.train[,1:ncol(dat.train)]
dat.train.y <- dat.train$FRACTURE

dat.train.y <- as.factor(as.character(dat.train.y))

#PCA
pc.result<-prcomp(dat.train.x,scale=TRUE)
pc.scores<-pc.result$x
pc.scores<-data.frame(pc.scores)
pc.scores$FRACTURE<-dat.train.y

PCA <- pc.result$rotation
PCA

#Scree plot
pc.eigen<-(pc.result$sdev)^2
pc.prop<-pc.eigen/sum(pc.eigen)
pc.cumprop<-cumsum(pc.prop)
plot(1:11,pc.prop,type="l",main="Scree Plot",ylim=c(0,1),xlab="PC #",ylab="Proportion of Variation")
 lines(1:11,pc.cumprop,lty=3)

#Use ggplot2 to plot the first few pc's
ggplot(data = pc.scores, aes(x = PC1, y = PC2)) +  geom_point(aes(col=FRACTURE), size=1)+ geom_hline(yintercept = 0, colour = "gray65") +  geom_vline(xintercept = 0, colour = "gray65") +  ggtitle("PCA plot of Osteo Study")

ggplot(data = pc.scores, aes(x = PC1, y = PC3)) +  geom_point(aes(col=FRACTURE), size=1)+ geom_hline(yintercept = 0, colour = "gray65") +  geom_vline(xintercept = 0, colour = "gray65") +  ggtitle("PCA plot of Osteo Study")

ggplot(data = pc.scores, aes(x = PC2, y = PC3)) +  geom_point(aes(col=FRACTURE), size=1)+ geom_hline(yintercept = 0, colour = "gray65") +  geom_vline(xintercept = 0, colour = "gray65") +  ggtitle("PCA plot of Osteo Study")
```
##Clustering

```{r fig.width=12}
#Lets look at a heatmap using hierarchical clustering to see if the 
#response naturually clusters out using the predictors
  
#Transposting the predictor matrix and giving the response categories its
#row names.
library(RColorBrewer)
x<-t(dat.train.x)
colnames(x)<-dat.train.y
pheatmap(x,annotation_col=data.frame(FRACTURE=dat.train.y),scale="row",legend=T,color=colorRampPalette(c("blue","white", "red"), space = "rgb")(100))

##logistic regression
dat.train.x <- as.matrix(dat.train.x)
library(glmnet)
cvfit <- cv.glmnet(dat.train.x, dat.train.y, family = "binomial", type.measure = "class", nlambda = 1000)
plot(cvfit)
coef(cvfit, s = "lambda.min")
```
##Creating a balanced dataset by Oversampling and check accuracy.
<p>Below we created a balanced dataset using ROSE package. The idea is that by creating an over sampled dataset with randomly duplicated records, we prodcued a more balanced dataset of subjects.
```{r}
library(ROSE)
#Look at the proporation of subjecs with Fractures from the Full dataset.
table(set_noID$FRACTURE)
prop.table(table(set_noID$FRACTURE))

#Create an oversampled dataset from Training dataset
over <- ovun.sample(FRACTURE~.,data=trainingData,method="over", N=564)$data
table(over$FRACTURE)

prop.table(table(over$FRACTURE))

#fit a logistic regressio to unblanced training set
fit.dataset <- glm(formula=FRACTURE~ ., data = over, family="binomial")
pred.fit.dataset <- predict(fit.dataset, newdata = over, type="response")
#Check Accuracy of fitted model.
accuracy.meas(over$FRACTURE,pred.fit.dataset, threshold=.05)

#Check Accuracy of Test dataset using ROC curve
roc.curve(over$FRACTURE, pred.fit.dataset, plotit = TRUE)
```
```