---
title: "test_interaction_data"
output:
  html_document:
    toc: yes
    keep_md: yes
  pdf_document:
    toc: yes
    toc_depth: 2
date: "August 10, 2018"
---

```{r setup_interaction0, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(knitr)
library(kableExtra)
library("ISLR")
library("SmartEDA")


data_setup_script <- here::here("code", "data_functions.R")
source(data_setup_script)
```


### Get Glow dataset

```{r get_dataset}

glow <- read_glow_dataset()

```

### model interactions - main effects

```
model_z1 <- glm(FRACTURE ~ AGE, family = binomial, data = glow)
model_z2 <- glm(FRACTURE ~ WEIGHT, family = binomial, data = glow)
model_z3 <- glm(FRACTURE ~ HEIGHT, family = binomial, data = glow)
model_z4 <- glm(FRACTURE ~ BMI, family = binomial, data = glow)
model_z5 <- glm(FRACTURE ~ PRIORFRAC, family = binomial, data = glow)
model_z6 <- glm(FRACTURE ~ PREMENO, family = binomial, data = glow)
model_z7 <- glm(FRACTURE ~ MOMFRAC, family = binomial, data = glow)
model_z8 <- glm(FRACTURE ~ ARMASSIST, family = binomial, data = glow)
model_z9 <- glm(FRACTURE ~ SMOKE, family = binomial, data = glow)
model_z10 <- glm(FRACTURE ~ RATERISK, family = binomial, data = glow)

## AGE          0.05289    0.01163   4.548 5.42e-06 ***
## WEIGHT      -0.005197   0.006415 -0.810    0.418
## HEIGHT      -0.05167    0.01709  -3.022  0.00251 **
## BMI          0.005758   0.017185  0.335  0.73760  
## PRIORFRACYes 1.0638     0.2231    4.769 1.85e-06 ***
## PREMENOYes   0.05077    0.25921   0.196    0.845  
## MOMFRACYes   0.6605     0.2810    2.351   0.0187 *  
## ARMASSISTYes 0.7091     0.2098    3.381 0.000723 ***
## SMOKEYes    -0.3077     0.4358   -0.706     0.48    
## RATERISKSame      0.5462     0.2664   2.050   0.0404 *  
## RATERISKGreater   0.9091     0.2711   3.353   0.0008 ***

> code below:
```

This leads us to consider the covariates above that are significant in the univariate results above at the 25% level.

AGE, HEIGHT, PRIORFRAC, MOMFRAC, ARMASSIST, RATERISK {SAME, GREATER}

```{r model_full_and_summary}

# fit a univariate logistic regression model for each covariate
# continuous - AGE WEIGHT HEIGHT BMI
# categorical - PRIORFRAC PREMENO MOMFRAC ARMASSIST SMOKE RATERISK

# model0
#model_z1 <- glm(FRACTURE ~ AGE, family = binomial, data = glow)
#model_z2 <- glm(FRACTURE ~ WEIGHT, family = binomial, data = glow)
#model_z3 <- glm(FRACTURE ~ HEIGHT, family = binomial, data = glow)
#model_z4 <- glm(FRACTURE ~ BMI, family = binomial, data = glow)
#model_z5 <- glm(FRACTURE ~ PRIORFRAC, family = binomial, data = glow)
#model_z6 <- glm(FRACTURE ~ PREMENO, family = binomial, data = glow)
#model_z7 <- glm(FRACTURE ~ MOMFRAC, family = binomial, data = glow)
#model_z8 <- glm(FRACTURE ~ ARMASSIST, family = binomial, data = glow)
#model_z9 <- glm(FRACTURE ~ SMOKE, family = binomial, data = glow)
#model_z10 <- glm(FRACTURE ~ RATERISK, family = binomial, data = glow)

#summary(model_z1)
#summary(model_z2)
#summary(model_z3)
#summary(model_z4)
#summary(model_z5)
#summary(model_z6)
#summary(model_z7)
#summary(model_z8)
#summary(model_z9)
#summary(model_z10)

# not interesting due to all variables (i.e. SUB_ID, SITE_ID, PHY_ID)
# model00 <- glm(FRACTURE ~ ., family = binomial, data = glow)
# summary(model00)

# full model, order by continuous, then factor
model0 <- glm(FRACTURE ~ AGE + WEIGHT + HEIGHT + BMI + PRIORFRAC + PREMENO + MOMFRAC + ARMASSIST + SMOKE + RATERISK, family = binomial, data = glow) 
summary(model0)

# fit model # note - should remove below model1
model0_fitted <- update(model0, . ~ . - WEIGHT - BMI - PREMENO - SMOKE)
summary(model0_fitted)

# build model with following covariates (drop WEIGHT, BMI, PREMENO, SMOKE)
# AGE, HEIGHT, PRIORFRAC, MOMFRAC, ARMASSIST, RATERISK {SAME, GREATER}

model1 <- glm(FRACTURE ~ AGE + HEIGHT + PRIORFRAC + MOMFRAC + ARMASSIST + RATERISK, family = binomial, data = glow) 
summary(model1)

# from above result, adding back the removed covariates we see they are not needed to keep the remaining covariates
# this becomes the model, adding back removed covariates WEIGHT, BMI, PREMENO, SMOKE the coefficients didn't become significant
# this becomes the main effects model


# need to check scale of logit for remaining continous variables AGE HEIGHT
# assume HEIGHT is linearin logit

```

## The main effects model
model1 <- glm(FRACTURE ~ AGE + HEIGHT + PRIORFRAC + MOMFRAC + ARMASSIST + RATERISK, family = binomial, data = glow) 

use lrtest from package lmtest


### test interactions for the following:

5. AGE: [HEIGHT, PRIORFRAC, MOMFRAC, ARMASSIST, RATERISK]
4. HEIGHT: [PRIORFRAC, MOMFRAC, ARMASSIST, RATERISK]
3. PRIORFRAC: [MOMFRAC, ARMASSIST, RATERISK]
2. MOMFRAC: [ARMASSIST, RATERISK]
1. ARMASSIST: RATERISK  

total 15 interactions

```{r loglikelihood_testing}
library(lmtest)

# model AGE* , HEIGHT* , PRIORFRAC*

model_effects <- glm(FRACTURE ~ AGE + HEIGHT + PRIORFRAC + MOMFRAC + ARMASSIST + RATERISK, family = binomial, data = glow) 
lrtest(model_effects)

# (5) AGE: [HEIGHT, PRIORFRAC, MOMFRAC, ARMASSIST, RATERISK]

test <- model_effects
test <- update(test, . ~ . + AGE:HEIGHT)
summary(test)
lrtest(test)

test <- model_effects
test <- update(test, . ~ . + AGE:PRIORFRAC)
summary(test)
lrtest(test)

test <- model_effects
test <- update(test, . ~ . + AGE:MOMFRAC)
summary(test)
lrtest(test)

test <- model_effects
test <- update(test, . ~ . + AGE:ARMASSIST)
summary(test)
lrtest(test)

test <- model_effects
test <- update(test, . ~ . + AGE:RATERISK)
summary(test)
lrtest(test)

# (4) HEIGHT: [PRIORFRAC, MOMFRAC, ARMASSIST, RATERISK]

test <- model_effects
test <- update(test, . ~ . + HEIGHT:PRIORFRAC)
summary(test)
lrtest(test)

test <- model_effects
test <- update(test, . ~ . + HEIGHT:MOMFRAC)
summary(test)
lrtest(test)

test <- model_effects
test <- update(test, . ~ . + HEIGHT:ARMASSIST)
summary(test)
lrtest(test)

test <- model_effects
test <- update(test, . ~ . + HEIGHT:RATERISK)
summary(test)
lrtest(test)

# (3) PRIORFRAC: [MOMFRAC, ARMASSIST, RATERISK]

test <- model_effects
test <- update(test, . ~ . + PRIORFRAC:MOMFRAC)
summary(test)
lrtest(test)

test <- model_effects
test <- update(test, . ~ . + PRIORFRAC:ARMASSIST)
summary(test)
lrtest(test)

test <- model_effects
test <- update(test, . ~ . + PRIORFRAC:RATERISK)
summary(test)
lrtest(test)

# (2) MOMFRAC: [ARMASSIST, RATERISK]

test <- model_effects
test <- update(test, . ~ . + MOMFRAC:ARMASSIST)
summary(test)
lrtest(test)

test <- model_effects
test <- update(test, . ~ . + MOMFRAC:RATERISK)
summary(test)
lrtest(test)

# (1) ARMASSIST: RATERISK  

test <- model_effects
test <- update(test, . ~ . + ARMASSIST:RATERISK)
summary(test)
lrtest(test)


```

### Results from interactions
``` 
## AGE:HEIGHT           0.001065   0.002095  0.508  0.6113  
   AGE:PRIORFRACYes    -0.05864    0.02583  -2.270  0.023188 *  
## AGE:MOMFRACYes      -0.01353    0.03264  -0.414  0.67854   
## AGE:ARMASSISTYes    -0.01132    0.02479  -0.457  0.64802   
## AGE:RATERISKSame    -0.03999    0.03151  -1.269  0.20438   
## AGE:RATERISKGreater -0.05021    0.03202  -1.568  0.11690  
## HEIGHT:PRIORFRACYes  0.01915    0.03770   0.508  0.61146   
## HEIGHT:MOMFRACYes    0.07401    0.04718   1.569  0.11675   
## HEIGHT:ARMASSISTYes -0.04419    0.03594  -1.229  0.21890   
## HEIGHT:RATERISKSame      -0.01577    0.04550  -0.347  0.72890   
## HEIGHT:RATERISKGreater    0.02394    0.04528   0.529  0.59695 
   PRIORFRACYes:MOMFRACYes  -1.07823    0.65021  -1.658  0.09726 . 
## PRIORFRACYes:ARMASSISTYes 0.24587    0.48467   0.507  0.61194   
## PRIORFRACYes:RATERISKSame     0.001597   0.625563   0.003  0.99796   
## PRIORFRACYes:RATERISKGreater  0.208811   0.624586   0.334  0.73814   
   MOMFRACYes:ARMASSISTYes  -1.33817    0.62405  -2.144  0.03201 * 
## MOMFRACYes:RATERISKSame     1.04615    0.95957   1.090  0.27561   
## MOMFRACYes:RATERISKGreater  0.36775    0.96207   0.382  0.70227   
## ARMASSISTYes:RATERISKSame     0.10760    0.56723   0.190  0.84956   
## ARMASSISTYes:RATERISKGreater -0.60953    0.58200  -1.047  0.29496   

```

## Add to main effects model

### we find three interactions, AGE:PRIORFRACYes, PRIORFRACYes:MOMFRACYes, MOMFRACYes:ARMASSISTYes

```{r fit_new_model}

model_effects_new <- glm(FRACTURE ~ AGE + HEIGHT + PRIORFRAC + MOMFRAC + ARMASSIST + RATERISK + 
                       AGE:PRIORFRAC + PRIORFRAC:MOMFRAC + MOMFRAC:ARMASSIST, family = binomial, data = glow) 
summary(model_effects_new)


# create final model with interactions terms AGE:PRIORFRAC + MOMFRAC:ARMASSIST
model_effects_final <- glm(FRACTURE ~ AGE + HEIGHT + PRIORFRAC + MOMFRAC + ARMASSIST + RATERISK + 
                       AGE:PRIORFRAC + MOMFRAC:ARMASSIST, family = binomial, data = glow) 
summary(model_effects_final)


```

## Final Interaction Model


```
(Intercept)
AGE                      0.05890    0.01666   3.535 0.000408 ***
HEIGHT                  -0.04413    0.01848  -2.388 0.016949 *  
PRIORFRACYes             4.65073    1.88342   2.469 0.013538 *  
MOMFRACYes               1.19902    0.39487   3.036 0.002393 ** 
ARMASSISTYes             0.61423    0.25358   2.422 0.015426 *  
RATERISKGreater          0.72116    0.29660   2.431 0.015040 *  
AGE:PRIORFRACYes        -0.05610    0.02600  -2.158 0.030950 *  
MOMFRACYes:ARMASSISTYes -1.26534    0.62377  -2.029 0.042507 *  

FRACTURE ~ AGE + HEIGHT + PRIORFRAC + MOMFRAC + ARMASSIST + RATERISK + AGE:PRIORFRAC + MOMFRAC:ARMASSIST

```


```{r interaction_model_ROC}

library(pROC)
library(vcdExtra)

# vcov(model_effects_final)

HLtest(model_effects_final)

glow$predict_mfinal <- predict(model_effects_final, type = "response")
with(glow, addmargins(table(glow$predict_mfinal > 0.5, glow$FRACTURE)))

(roc_final_model <- roc(glow$FRACTURE ~ glow$predict_mfinal, data = glow))
plot(roc_final_model, main = "ROC: Final Effects Model")

```

misc

```{r additional_interaction_model_ROC}

## restart with clean data
glow <- read_glow_dataset()

model_last <- glm(FRACTURE ~ AGE:PRIORFRAC + HEIGHT + MOMFRAC:ARMASSIST + I(as.integer(RATERISK) == 3), family = binomial, data = glow)
HLtest(model_last)
summary(HLtest(model_last))

# classification table
glow$predict_last <- predict(model_last, type = "response")
with(glow, addmargins(table(predict_last > 0.5, FRACTURE)))

# Sensitivy, specificity, ROC (using pROC)
roc_model_last <- roc(glow$FRACTURE ~ glow$predict_last, data = glow)
plot(roc_model_last, main = "ROC: Last model")

# create table 
vars <- c("thresholds","sensitivities","specificities")
model_table <- data.frame(roc_model_last[vars])

findIndex <- function(x, y) which.min( (x-y)^2 )
cutPoints <- seq(0.05, 0.75, by = 0.05)

tableIndex <- mapply(findIndex, y = cutPoints, MoreArgs = list(x = roc_model_last$thresholds))

model_table[tableIndex, ]

# plot
plot(specificities ~ thresholds, xlim = c(0, 1), type = "l",
xlab = "probability cutoff", ylab = "sensitivity / specificity",
ylim = c(0, 1), data = model_table, main = "probability sensitivity")
with(model_table, lines(thresholds, sensitivities, col = "red"))
legend(x = 0.75, y = 0.55, legend = c("Sensitivity", "Specificity"),
lty = 1, col = c("red","black"))
abline(h = c(0, 1), col = "grey80", lty = "dotted")

```



